---
title: "Chapter2_ISLR"
author: "Avi Sooriyarachchi"
date: "3/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the necessary libraries
```{r}
library(MASS)
library(ISLR)
```

check out the Boston data set from the MASS library
```{r}
View(Boston)
names(Boston)
```

If you need to learn more about the dataset do the following
```{r}
?Boston
```

Do linear regression with medv as Y and lstat as X
```{r}
lm.fit = lm(medv~lstat, data =Boston)
attach(Boston)
lm.fit = lm(medv~lstat)
```

See what was calculated for the linear regression fit (Ordinary Least Squares)
```{r}
lm.fit
```



This isn't very informative, it is helpful to get a summary as shown below 
```{r}
summary(lm.fit)
```


Use names to figure out what else is computed in lm.fit
```{r}
names(lm.fit)
```

```{r}
coef(lm.fit)
```


Finding a confidence interval for the coefficients of the linear fit, perform the following 
```{r}
confint(lm.fit)
```

To find confidence intervals for a given value of lstat
```{r}
attach(Boston)
predict(lm.fit, data.frame(lstat=c(5,10,15)), interval = "confidence")
```

```{r}
predict (lm.fit ,data.frame(lstat=c(5 ,10 ,15) ),interval ="prediction")
```

We can explore different ways of changing the appearance of the regression plot 
```{r}
plot(lstat, medv)
abline(lm.fit)
abline(lm.fit, lwd = 3)
abline(lm.fit, lwd = 3, col = "red")
plot(lstat, medv, col = "red")
plot(lstat, medv, pch = 20)


```


Getting all 4 diagnostic plot at once 
```{r}
par(mfrow = c(2,2))
plot(lm.fit)

```

Getting diagnostic plot using residuals function 
```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```
Computing Leverage statistics
```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```


Fitting a multiple regrassion line to a given set of data
```{r}
lm.fit  = lm(medv~lstat + age, data=Boston)
summary(lm.fit)
```

In order to consider all independent variable
```{r}
lm.fit = lm(medv~., data = Boston)
summary(lm.fit)
```

To examine the presence of multicollinearity in this data, computing VIF is useful. To use this we have to use vif function given in the car package 
```{r}
library(car)
vif(lm.fit)
```

Running a multiple regression excluding a certain factor
```{r}
lm.fit1 = lm(medv~.-age, data = Boston)
summary(lm.fit1)
```



Alternatively, the update() function can be used 
```{r}
lm.fit1 = update(lm.fit, ~.-age)
summary(lm.fit1)
```

Interaction terms 
lstat:black is an interaction term between lstat and black.
However a better way to do this lstat*black <==> lstat + black + lstat:black
```{r}
summary(lm(medv~black*lstat, data = Boston))
```

Wrap the non linear entity in a () in the R formula. An escape chatacter of sorts in this context
```{r}
lm.fit2 = lm(medv ~lstat + I(lstat^2))
summary(lm.fit2)
```

The anova() function performs a hypothesis
test comparing the two models. The null hypothesis is that the two models
fit the data equally well, and the alternative hypothesis is that the full
model is superior.
Use anova() function to show how better the quadratic fit is in comparison to the linear fit. 
```{r}
lm.fit = lm(medv~lstat)
anova(lm.fit, lm.fit2)
```


A better approach involves using the poly() function
poly() to create the polynomial within lm(). For example, the following command
produces a fifth-order polynomial fit


```{r}
lm.fit5 = lm(medv~poly(lstat, 5))
summary(lm.fit5)
```


In addition to polynomial trasformations we can use log transformations of the predictor variables
```{r}
summary(lm(medv~log(rm), data=Boston))
```

```{r}
library(ISLR)
```

Qualitative predictor variables using the Carseats dataset in R 
```{r}
?Carseats
```


```{r}

remove(list = ls())

```
```{r}
View(Carseats)
```

```{r}
attach(Carseats)
```

```{r}
names(Carseats)
```

R qutomtically generates dummy variables for qualitative variables 
```{r}
lm.fit = lm(Sales~.+Income:Advertising+Price:Age, data = Carseats)
summary(lm.fit)
```

```{r}
?contrasts
```

contrasts() function returns the coding that R uses for the dummy variables
```{r}
attach(Carseats)
contrasts(ShelveLoc)
```

Writing functions in R
Let's try to create a Library names LoadLibraries
```{r}
LoadLibraries
LoadLibraries()
```

```{r}
LoadLibraries <- function(){
  library(ISLR)
  library(MASS)
  print("The libraries have loaded ")
}
```

```{r}
LoadLibraries()
```

